---
title: "Particle Identification ML"
author: "Gerhard Viljoen"
date: "12/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Read in all the json files, created from python dictionaries

```{r}
# rm(list=ls())
# 
# require(jsonlite)
# require(readtext)
# 
# #PDG codes
# pdg.elec <- c(11,-11)
# pdg.pion <- c(-211,211)
# 
# files <- list.files(path="~/Thesis data/SemiFullData", pattern="*json", full.names=T, recursive=FALSE)
# 
# j <- fromJSON(files[1])
# 
# for(i in 2:length(files)){
# 
#   f <- fromJSON(files[i])
#   j <- c(j,f)
# }
# 
# length(j)
# 
# save(j,file="~/Thesis data/SemiFullData/fulljson.rdata")
```

#Threshold pad data

Load full json record of all extracted events:

```{r}
# rm(list=ls())
load("~/Thesis data/SemiFullData/fulljson.rdata")
```

Extract pad data

```{r}

#pads <- sapply(j,`[[`,"layer0")

```

Replace all pads that have data, with only those bin-rows which are all non-zero

```{r}
#no.zero.row.pads <- list()
# 
# for(i in 1:length(pads)){
#   if(is.null(pads[[i]])){
#     no.zero.row.pads[[i]] <- NULL
#   }else if(typeof(pads[[i]])=="list"){
#     no.zero.row.pads[[i]] <- NULL
#   }else{
#   this.pad <- as.matrix(pads[[i]])
#   r <- rowSums(this.pad)
#   r <- which(r==0)
#   
#   if(length(r)==11){
#     this.pad <- NULL
#   }else{
#     this.pad <- this.pad[-r,]
#   }
#   
#   
#   
#   no.zero.row.pads[[i]] <- this.pad
#   }
# }
# 
# save(no.zero.row.pads,file="~/Thesis data/SemiFullData/no_zero_row_pads.rdata")
```


Create look-up tables to find pad data for both electrons and pions, and add an indicator column for which entries have no pad data.

Also create an indicator of which tracks' trackIDs are 0, since this may indicate that an error has occurred


```{r}
# tracks <- sapply(j, `[[`,"track")
# tracks <- as.numeric(tracks)
# 
# 
# pdg <- sapply(j, `[[`,"pdgCode")
# pdg <- as.numeric(pdg)
# 
# exclude <- rep(0,length(no.zero.row.pads))
# 
# for(i in 1:length(no.zero.row.pads)){
#   if(is.null(no.zero.row.pads[[i]])){
#     exclude[i] <- 1
#   }
# }
# 
# exclude2 <- rep(0,length(no.zero.row.pads))
# 
# for(i in 1:length(tracks)){
#   if(tracks[i]==0){
#     exclude2[i] <- 1
#   }
# }
# 
# 
# look.up.table <- data.frame(cbind(pdg,1:length(pdg),exclude,exclude2))
# names(look.up.table) <- c("pdg","index","exclude.null.pads","exclude.track.id.0")
```

```{r}
# require(dplyr)
# pdg.elec <- c(11,-11)
# pdg.pion <- c(211,-211)
# electrons <- look.up.table[which(look.up.table$pdg %in% pdg.elec),]
# pions <- look.up.table[which(look.up.table$pdg %in% pdg.pion),]
# 
# electrons <- unique(electrons)
# pions <- unique(pions)
# 
# electrons <- electrons %>%
#   subset(exclude.null.pads==0) %>%
#   subset(exclude.track.id.0==0)
# 
# pions <- pions %>%
#   subset(exclude.null.pads==0)  %>%
#   subset(exclude.track.id.0==0)
# 
# electrons <- as.data.frame(electrons)
# pions <- as.data.frame(pions)
# 
# electrons <- electrons$index
# pions <- pions$index

```

Clear up some memory

```{r}

# rm(look.up.table)
# rm(j)
# rm(exclude)
# rm(exclude2)
# rm(tracks)
# rm(r)
# rm(this.pad)
# rm(i)
# rm(pdg)

# save(electrons,file="~/Thesis data/SemiFullData/electrons.rdata")
# save(pions,file="~/Thesis data/SemiFullData/pions.rdata")
# save(pads,file="~/Thesis data/SemiFullData/pads.rdata")
```

#Descriptive statistics

Get total energy deposition per pad, for pions and electrons, respectively:

```{r}
# electron.pad.sum <- c()
#   
# for(i in electrons){
#   this.pad.sum <- sum(as.numeric(pads[[i]]))
#   electron.pad.sum <- c(electron.pad.sum,this.pad.sum)
# }
# 
# pion.pad.sum <- c()
# 
# for(i in pions){
#   this.pad.sum <- sum(as.numeric(pads[[i]]))
#   pion.pad.sum <- c(pion.pad.sum,this.pad.sum)
# }

```

```{r}
# electrons <- data.frame(cbind(electrons,electron.pad.sum))
# pions <- data.frame(cbind(pions,pion.pad.sum))
# 
# names(electrons) <- c("index","total.charge.deposit")
# names(pions) <- c("index","total.charge.deposit")
```

Get 5 number summary statistics, for each pad where 0-sum bin-rows have been removed.


```{r}
# electron.pad.summary <- numeric(6)
# 
# for(i in electrons$index){
#   this.pad <- summary(as.numeric(no.zero.row.pads[[i]]))
#   electron.pad.summary <- rbind(electron.pad.summary,this.pad)
# }
# 
# electron.pad.summary <- as.data.frame(electron.pad.summary)
# names(electron.pad.summary) <- c("Min","1Q","Median","Mean","3Q","Max")
# 
# pion.pad.summary <- numeric(6)
# 
# for(i in pions$index){
#   this.pad <- summary(as.numeric(no.zero.row.pads[[i]]))
#   pion.pad.summary <- rbind(pion.pad.summary,this.pad)
# }
# 
# pion.pad.summary <- as.data.frame(pion.pad.summary)
# names(pion.pad.summary) <- c("Min","1Q","Median","Mean","3Q","Max")
```

Clean up memory:

```{r}
# electron.pad.summary <- electron.pad.summary[-1,]
# pion.pad.summary <- pion.pad.summary[-1,]
# 
# electrons <- data.frame(cbind(electrons,electron.pad.summary))
# pions <- data.frame(cbind(pions,pion.pad.summary))
# 
# rm(pion.pad.summary)
# rm(electron.pad.summary)
# rm(electron.pad.sum)
# rm(pion.pad.sum)
# 
# 
# save(electrons,file="~/Thesis data/SemiFullData/electrons.rdata")
# save(pions,file="~/Thesis data/SemiFullData/pions.rdata")
```

Get the number of non-zero bin-rows per detector pad, for electrons and pions:

```{r}
# electron.timebins <- c()
# 
# for(i in electrons$index){
# 
#   if(is.null(nrow(no.zero.row.pads[[i]]))){
#     this.num.bin <- 0
#   }else{
#     this.num.bin <- nrow(no.zero.row.pads[[i]])
#   }
# 
# 
#   electron.timebins <- c(electron.timebins,this.num.bin)
# 
# }
# 
# pion.timebins <- c()
# 
# for(i in pions$index){
# 
#     if(is.null(nrow(no.zero.row.pads[[i]]))){
#     this.num.bin <- 0
#   }else{
#   this.num.bin <- nrow(no.zero.row.pads[[i]])
#   
# 
#   }
#   
#   pion.timebins <- c(pion.timebins,this.num.bin)
# }
# 
# 
# electrons$num.bins <- electron.timebins
# pions$num.bins <- pion.timebins
# 
# save(electrons,file="~/Thesis data/SemiFullData/electrons.rdata")
# save(pions,file="~/Thesis data/SemiFullData/pions.rdata")
```

Clean up memory

```{r}
# rm(electron.timebins)
# rm(pion.timebins)
```


Get the column means for each pad, where non-zero time-bins have been removed:

```{r}
# electron.compressed.bins <- numeric(24)
# 
# for(i in electrons$index){
#   
#   if(is.null(dim(no.zero.row.pads[[i]]))){
#     this.entry <- as.numeric(no.zero.row.pads[[i]])
#   }else{
#   
# this.entry <- as.numeric(colSums(no.zero.row.pads[[i]]))
# 
# 
#   }
#   
#   electron.compressed.bins <- rbind(electron.compressed.bins,this.entry)
# }
# 
# pion.compressed.bins <- c()
# 
# for(i in pions$index){
#   
#     if(is.null(dim(no.zero.row.pads[[i]]))){
#     this.entry <- as.numeric(no.zero.row.pads[[i]])
#   }else{
#   
#   this.entry <- as.numeric(colSums(no.zero.row.pads[[i]]))
#   
#   }
#   pion.compressed.bins <- rbind(pion.compressed.bins,this.entry)
# }
# 
# electron.compressed.bins <- electron.compressed.bins[-1,]
# 
# pion.compressed.bins <- pion.compressed.bins[-1,]
# 
# electron.compressed.bins <- as.data.frame(electron.compressed.bins)
# pion.compressed.bins <- as.data.frame(pion.compressed.bins)
# 
# names <- paste0("c",1:24)
# 
# names(electron.compressed.bins) <- names
# names(pion.compressed.bins) <- names
# 
# electrons <- data.frame(cbind(electrons,electron.compressed.bins))
# pions <- data.frame(cbind(pions,pion.compressed.bins))
# 
# save(electrons,file="~/Thesis data/SemiFullData/electrons.rdata")
# save(pions,file="~/Thesis data/SemiFullData/pions.rdata")
# 
# electrons$id <- "electron"
# pions$id <- "pion"
# 
# electrons$id <- as.factor(electrons$id)
# pions$id <- as.factor(pions$id)
```

Clear up memory

```{r}
# rm(pion.compressed.bins)
# rm(electron.compressed.bins)
```

```{r}
# dat <- data.frame(rbind(electrons,pions))
# rm(electrons)
# rm(pions)
#rm(no.zero.row.pads)
# save(dat,file="~/Thesis data/SemiFullData/unscaled_data.rdata")
```

Scale data

```{r}
# dat[,2:33] <- scale(dat[,2:33])
# save(dat,file="~/Thesis data/SemiFullData/scaled_data.rdata")
# 

# load("~/Thesis data/SemiFullData/scaled_data.rdata")
# i <- dat$index
# rm(dat)
```

Unroll pad data into a vector for each entry:

```{r}
# pad.data <- matrix(nrow=length(pads),ncol=11*24)
# 
# for(j in i){
#   pad.data[j,] <- as.numeric(pads[[j]])
# }
# 
# pad.data <- pad.data[i,]
# 
# pad.data <- scale(pad.data)
# 
# rm(pads)
# rm(j)
# rm(i)

# load("~/Thesis data/SemiFullData/scaled_data.rdata")
# 
# dat <- data.frame(cbind(dat,pad.data))
# 
# dat <- dat[,c(1,34,2:33,35:298)]
```

```{r}
# rm(pad.data)
# require(knitr)
# save(dat,file="~/Thesis data/SemiFullData/final_data.rdata")
rm(list=ls())
load("~/Thesis data/SemiFullData/final_data.rdata")
na.omit(dat)
kable(head(dat,100))
```

```{r}
# any(is.na(dat))
# dat <- na.omit(dat)
# 
# rownames(dat) <- NULL
# 
# index <- dat$index
# 
# dat <- dat[,-1]
# 
# dat$id <- as.character(dat$id)
# 
# dat$electron <- ifelse(dat$id=="electron",1,0)
# 
# id <- dat$id
# 
# dat <- dat[,-1]
```

```{r}
id.index <- data.frame(cbind(id,index))

require(dplyr)

electron.id <- id.index[id.index$id=="electron",]

pion.id <- id.index[id.index$id=="pion",]

electron.train <- base::sample(electron.id$index)


```


#Particle Identification:

Simple linear regression:

```{r}
linmod1 <- lm(electron~.,data=dat)
```



































